XXX

### 2012-03-12 paper/engNERwiki.tex
mai megbeszeltek es otletek Nothman-bol beleirva

Wikipedia is recognized as one of the largest available collection of entities, and also as an external knowledge to improve the accuracy of NER. The most obvious utilization of Wikipedia for NER is extracting gazetteers containing person names, locations or organizations \cite{Toral:06}. Creating dictionaries of entities is also a common step of named entity disambiguation \cite{Bunescu:06,Cucerzan:07}. Nadeau et al. \shortcite{Nadeau:06} use similar lists in an unsupervised NER system, while Kazama and Torisawa \shortcite{Kazama:07} showed that automatic extraction of category labels from Wikipedia improves the accuracy of a supervised NE-tagger.

@@

Automatic creation of training data from Wikipedia for NER:
-classification of articles into CoNLL entity classes \cite{Nothman:08}

@@

@@

@@

@@

@@

@@

>>>>

Wikipedia is recognized as one of the largest available collection of entities, and also as an external knowledge to improve the accuracy of NER. The most obvious utilization of Wikipedia for NER is extracting gazetteers containing person names, locations or organizations \cite{Toral:06}. Creating dictionaries of entities is also a common step of named entity disambiguation \cite{Bunescu:06,Cucerzan:07}. Not only supervised systems use lists generated from Wikipedia: Nadeau et al. \shortcite{Nadeau:06} use similar lists in an unsupervised NER system.

Wikipedia's knowledge may be incorporated as features in NER learning, e.g. Kazama and Torisawa \shortcite{Kazama:07} showed that automatic extraction of category labels from Wikipedia improves the accuracy of a supervised NE-tagger.

A third approach to improve NER with Wikipedia is the automatic creation of training data. Richman and Schone \shortcite{Richman:08} built corpora for less commonly taught languages annotated with NE-tags. They used the inherent category structure of Wikipedia to determine the NE-type of a proposed entity. Nothman et al.~\shortcite{Nothman:08} used a similar method to create a NE-annotated text in English. They transformed the Wikipedia links into NE-annotations by classifying the target articles into standard entity classes. Their approach to classification is based primarily on category head nouns and definitional opening sentences of articles.

not open access XXX

más nyelvekre is csinálták XXX

domain-specificity: not really

Szeged NER korpusz kicsi, ahhoz képest ez nagy

DBPediát előrébb

@@

%Automatic creation of training data from Wikipedia for NER:
%-classification of articles into CoNLL entity classes \cite{Nothman:08}

@@

from \cite{Nothman:09}: Exceptions are made for common titlecase words, e.g. I, Mr., June, and sentence-initial words.

@@

In Wikipedia the nationalities (e.g. American and Australian) links to the corresponding places (America and Australia, respectively). However, in CoNLL scheme these types of entities take parts of MISC category. XXX Postprocessing: list look-up (nationalities, adjective forms of person who live in the corresponding country); if the name is POS-tagged as adjective, then it is not a NE

%nem biztos, hogy ide kell
Link boundaries need to be adjusted, e.g. to remove excess punctuation.

@@

tokenizálás: built in-house, tailored for Hungarian, because of the specialities of Hungarian (e.g. period after years in date expressions)

ocamorph + hundisambig, a lemmatizálás miatt

a KR-ban nincs jelölve a proper noun külön

DBPedia--CoNLL mapping for Hungarian: az angol oldalak vannak megcímkézve --> csak azok a magyar oldalak vannak letöltve és feldolgozva, amik megvannak angolul is
	--> future work: kibővíteni azokkal a magyar cikkekkel is, amik nincsenek angolul

filtering: az igés szűrő itt nem lesz

most van: 102 millió szó a disambiguation pages-zel együtt
az angol: 1,7 milliárd

according to the Szeged NER guideline

a nyelv és népnevek magyarul nem NE-k, vagyis nem kell az annotálásukkal foglalkozni

a tulajdonnevekhez kötőjellel kapcsolódó köznevek kérdésesek (pl. SAP-szoftvercég, Matáv-csoport)

a tulajdonnevekből képzett szavak sem NE-k

@@

measuring Wikipedia's inconsistency using n-gram tag variation (\cite{Nothman:09})
ez jó mérőszám lehet a gold standard korpuszok összehasonlítására
használhatjuk fordítva is: ha egy adott nucleus ugyanolyan n-gram környezetben nagyon sokszor van X-ként taggelve, és kevésszer Y-ként, akkor az Y-okat is átcsinálhatnánk X-szé

@@

az eredeti DBPedia osztályokat is megtarthatnánk, hiszen nem mindig célravezető a CoNLL osztályok használata, mivel azok elég tág kategóriák, amikbe olyan dolgok is tartoznak, amik nagyon eltérőek is lehetnek

később majd kirakjuk a scripteket is, és akár újra és újra frissíthető lesz

@@

\bibitem[\protect\citename{Nothman \bgroup et al.\egroup}2009]{Nothman:09}
\newblock 2009.
\newblock Analysing Wikipedia and Gold-Standard Corpora for NER Training.
\newblock In: {\em Proceedings of the 12th Conference of the European Chapter of the ACL}, pages 612--620.

\bibitem[\protect\citename{Richman and Schone}2008]{Richman:08}
Alexander E. Richman and Patrick Schone.
\newblock 2008.
\newblock Mining Wiki Resources for Multilingual Named Entity Recognition.
\newblock In: {\em Proceedings of ACL-08: HLT}, pages 1--9.
